# Summary Stats Automation

<!-- # ```{r} -->
<!-- # # Files to use later in rmd -->
<!-- # eiwg_files <- c("SST.csv", "Chla.csv", "beach.csv", "msddep.csv", "msdlat.csv") -->
<!-- # -->
<!-- # for(i in eiwg_files) { ##### BREAKS on GLPopulation.CSV #################### -->
<!-- # -->
<!-- #   name <- gsub(".csv","",i) -->
<!-- # -->
<!-- #   header_read<- read.csv(here::here("data-eiwg",i), nrows = 1) # grabs just the two rows at the top -->
<!-- #   title = colnames(header_read[1]) # title from column names -->
<!-- #   y_lab = as.character(header_read[1,][1]) #ylab from row 1 -->
<!-- # -->
<!-- # -->
<!-- #   dat<- assign(name, -->
<!-- #          read.csv(here::here("data-eiwg",i),skip = 2)) -->
<!-- #     colnames(dat) <- tolower(colnames(dat)) -->
<!-- #     dat<- dat %>% -->
<!-- #     tidyr::pivot_longer(!year, names_to = "region", values_to = "value") %>% -->
<!-- #     dplyr::mutate(title = c(title), # adds columns with title and y label for figures -->
<!-- #                    y_lab = c(y_lab)) -->
<!-- # -->
<!-- #   write.csv(dat, file = here::here(paste0("data-raw/", name, ".csv"))) #writes data to data-raw to be pull directly from functions below -->
<!-- # } -->
<!-- # -->
<!-- # -->
<!-- # ``` -->


```{r setup, include=FALSE, warning=FALSE, message=TRUE, echo=FALSE}

library(magrittr)

shade.alpha <- 0.3
shade.fill <- "lightgrey"
lwd <- 1
pcex <- 2
trend.alpha <- 0.5
trend.size <- 2
hline.size <- 1
hline.alpha <- 0.35
hline.lty <- "dashed"
label.size <- 5
hjust.label <- 1.5
letter_size <- 4
ylim_re <- c(5e6, 30e6)
ylim_rd <- c(1.75,2.75)
ylim_ra  <- c(0, 2e6)

#issues: Can't handle 5+ NAs, doesn't do monthly
summarize_me_for_fun<- function(data, time_step = "annual"){
  
  if(time_step == "annual"){
  name <- gsub(".csv","",i)
  header_read<- read.csv(here::here("data-eiwg",paste0(i, ".csv")), nrows = 1) # grabs just the two rows at the top
  title = colnames(header_read[1]) # title from column names
  y_lab = as.character(header_read[1,][1]) #ylab from row 1 
  

  dat<- assign(name,
         read.csv(here::here("data-eiwg",paste0(i, ".csv")), skip = 2))
  colnames(dat) <- tolower(colnames(dat))
  dat_all<- dat %>% 
    tidyr::pivot_longer(!year, names_to = "region", values_to = "value") %>% 
    dplyr::mutate(title = c(title), # adds columns with title and y label for figures
                   y_lab = c(y_lab))
  
  dat_all2 <- dat_all %>% 
    dplyr::group_by(region, title, y_lab) %>% 
    dplyr::summarise(mn = mean(value, na.rm = TRUE)) 
  
  sum_dat <- dat_all %>% 
    dplyr::group_by(region) %>% 
    dplyr::filter(year %in% c(max(year):(max(year) - 4))) %>% 
    dplyr::mutate(mn5 = mean(value, na.rm=TRUE)) %>% 
    dplyr::left_join(dat_all2, by = c("region", "title", "y_lab")) %>% 
    dplyr::summarise(status = dplyr::case_when(mn5>mn ~ "greater than", 
                                               mn5<mn ~ "less than"), 
                     mn5 = mn5)
  
  summary_df<- dat_all %>% 
    dplyr::group_by(region) %>% 
    dplyr::left_join(sum_dat, by = c("region")) %>%
    dplyr::mutate(#mn5 = mean(value, na.rm=TRUE),
                  delta = stats::predict(lm(year~value))[length(stats::predict(lm(year~value)))] - stats::predict(lm(year~value))[1], 
                  Z = abs(delta)-(sd(value, na.rm=TRUE)), 
                  trend = ifelse(Z > 0, paste0("significantly"), paste0("")), 
                  gauge = round(stats::ecdf(value)(mn5) *100)) %>% 
    
    dplyr::left_join(dat_all2, by = c("region", "title" ,"y_lab")) %>% 
    dplyr::summarise(indicator = unique(title), 
                     region = unique(region), 
                     #year = unique(year),
                     mn = unique(mn), 
                     mn5 = unique(mn5),
                     status = unique(status), 
                     trend = unique(trend), 
                     delta = unique(delta), 
                     Z = unique(Z),
                     gauge = unique(gauge))  
    
  }
  if(time_step == "monthly"){
    
    name <- gsub(".csv","",i)
    header_read<- read.csv(here::here("data-eiwg", paste0(i, ".csv"))) %>% 
      janitor::row_to_names(1) 
    y_lab = as.character(header_read[1,][2])
   
    dat_all<- read.csv(here::here("data-eiwg", paste0(i, ".csv"))) %>% 
      janitor::row_to_names(2) %>% 
      dplyr::rename_with(
        ~ dplyr::case_when(
            . == "Date" ~ "Year",
            . == "year" ~ "Year",
            TRUE ~ .)) %>% 
      dplyr::mutate(new_date = 1:length(Year)) %>% 
      dplyr::rename(value = 2) 

  dat_all2 <- dat_all%>% 
    dplyr::mutate(mn = mean(as.numeric(value, na.rm = TRUE)))
  
  sum_dat <- dat_all  %>% 
    dplyr::filter(new_date %in% c(max(new_date):(max(new_date) - 59))) %>% 
    dplyr::mutate(mn5 = mean(as.numeric(value, na.rm=TRUE))) %>% 
    dplyr::left_join(dat_all2, by = c("Year", "new_date", "value")) %>% 
    dplyr::mutate(status = dplyr::case_when(mn5>mn ~ "greater than", 
                                               mn5<mn ~ "less than"), 
                  mn5 = mn5)
  
  summary_df<- sum_dat %>% 
    dplyr::mutate(delta = stats::predict(lm(new_date~value))[length(stats::predict(lm(new_date~value)))],
                  stats::predict(lm(new_date~value))[1], 
                  Z = abs(delta)-(sd(value, na.rm=TRUE)), 
                  trend = ifelse(Z > 0, paste0("significantly"), paste0("")), 
                  gauge = round(stats::ecdf(value)(mn5) *100)) %>% 
    
    #dplyr::left_join(dat_all2, by = c("region", "title" ,"y_lab")) %>% 
    dplyr::summarise(indicator = y_lab,
                     #region = unique(region), 
                     #year = unique(year),
                     mn = unique(mn), 
                     mn5 = unique(mn5),
                     status = unique(status), 
                     trend = unique(trend), 
                     delta = unique(delta), 
                     Z = unique(Z),
                     gauge = unique(gauge))  
  
  }
  return(summary_df)
}
#issues: Can't handle 5+ NAs, doesn't do monthly

plot_me_for_fun<- function(data, time_step = "annual", save = TRUE){
  
  if(time_step == "annual"){
    name <- gsub(".csv","",i)
    header_read<- read.csv(here::here("data-eiwg",paste0(i, ".csv")), nrows = 1) # grabs just the two rows at the top
    title = colnames(header_read[1]) # title from column names
    y_lab = as.character(header_read[1,][1]) #ylab from row 1 
  
  

  dat<- assign(name,
         read.csv(here::here("data-eiwg",paste0(i, ".csv")), skip = 2))
  colnames(dat) <- tolower(colnames(dat))
  dat_all<- dat %>% 
    tidyr::pivot_longer(!year, names_to = "region", values_to = "value") %>% 
    dplyr::mutate(title = c(title), # adds columns with title and y label for figures
                   y_lab = c(y_lab))
  #dat <- read.csv(here::here("data-eiwg", paste0(i, ".csv")))
    
    values<- dat_all %>% 
      dplyr::group_by(region) %>% 
      dplyr::mutate(mn  = mean(value), 
                    upper = mn + sd(value), 
                    lower = mn - sd(value))
    
    pt<-dat_all %>% 
     dplyr::left_join(values, by = c("region", "value", "title", "y_lab", "year")) %>% 
      ggplot2::ggplot(ggplot2::aes(x = year, y = value, group = region))+
      ggplot2::annotate("rect", fill = "green", alpha = 0.3,
                        xmin = max(dat$year - 5), xmax = max(dat$year),
                        ymin = -Inf, ymax = Inf)+
      ggplot2::theme(panel.background = ggplot2::element_rect(fill = "transparent"), # bg of the panel
            panel.border = ggplot2::element_rect(colour = "black", fill=NA, size=1)) +
      ggplot2::geom_line(color = "red")+
      ggplot2::geom_point(color = "red")+
      ggplot2::geom_hline(ggplot2::aes(yintercept = mn),
                 size = hline.size,
                 alpha = hline.alpha,
                 linetype = hline.lty) +
      #geom_hline(aes(yintercept = sd_TEMP,),
      ggplot2::geom_hline(ggplot2::aes(yintercept = upper),
                 size = hline.size,
                 alpha = hline.alpha,
                 linetype = "dotted") +
      #geom_hline(aes(yintercept = -sd_TEMP,),
      ggplot2::geom_hline(ggplot2::aes(yintercept = lower),
                 size = hline.size,
                 alpha = hline.alpha,
                 linetype = "dotted") +
      ggplot2::ggtitle(paste(values$title))+
      ggplot2::ylab(paste(values$y_lab))+ ## Fix this with Will
      ggplot2::xlab(expression("Year"))+
      ggplot2::facet_wrap(~region, scales = "free")
  }
  
  
  if(time_step == "monthly"){
    name <- gsub(".csv","",i)
    header_read<- read.csv(here::here("data-eiwg", paste0(i, ".csv"))) %>% 
      janitor::row_to_names(1) 
    y_lab = as.character(header_read[1,][2])
    
    dat_all<- read.csv(here::here("data-eiwg", paste0(i, ".csv"))) %>% 
      janitor::row_to_names(2) %>% 
      dplyr::rename_with(
        ~ dplyr::case_when(
            . == "Date" ~ "Year",
            . == "year" ~ "Year",
            TRUE ~ .)) %>% 
      dplyr::mutate(new_date = 1:length(Year)) %>% 
      dplyr::rename(value = 2) %>% 
      dplyr::mutate(value = as.numeric(value))
    
    values<- dat_all %>% 
      dplyr::mutate(mn  = mean(value), 
                    upper = mn + sd(value), 
                    lower = mn - sd(value))
    
    pt<-dat_all %>% 
     #dplyr::left_join(values, by = c("value", "title", "y_lab", "new_date")) %>% 
      ggplot2::ggplot(ggplot2::aes(x = new_date, y = value))+
      ggplot2::annotate("rect", fill = "green", alpha = 0.3,
                        xmin = max(dat_all$new_date - 60), xmax = max(dat_all$new_date),
                        ymin = -Inf, ymax = Inf)+
      ggplot2::theme(panel.background = ggplot2::element_rect(fill = "transparent"), # bg of the panel
            panel.border = ggplot2::element_rect(colour = "black", fill=NA, size=1)) +
      ggplot2::geom_line(color = "red")+
      ggplot2::geom_point(color = "red")+
      ggplot2::geom_hline(ggplot2::aes(yintercept = values$mn),
                 size = hline.size,
                 alpha = hline.alpha,
                 linetype = hline.lty) +
      #geom_hline(aes(yintercept = sd_TEMP,),
      ggplot2::geom_hline(ggplot2::aes(yintercept = values$upper),
                 size = hline.size,
                 alpha = hline.alpha,
                 linetype = "dotted") +
      #geom_hline(aes(yintercept = -sd_TEMP,),
      ggplot2::geom_hline(ggplot2::aes(yintercept = values$lower),
                 size = hline.size,
                 alpha = hline.alpha,
                 linetype = "dotted") +
      ggplot2::ggtitle(paste(dat_all$y_lab))+
      ggplot2::ylab(paste(y_lab))+ ## Fix this with Will
      ggplot2::xlab(expression("Time"))#+
      #ggplot2::facet_wrap(~region, scales = "free")
  }
    
    print(pt)
  #   if(save == TRUE){
  #     png(here::here(paste0("plots_saved/", something, ".png")))
  # }else{ }
}

# plain_language<- function(data){
#   plain<-  summarydf %>% 
#     group_by(region, title) %>% 
#     dplyr::mutate(PlainLanguage = c(paste("this figure shows", title, " in ", y_lab, "ranging from ", min(value), "to", max(value), 
#                                           "from", min(year), "to ", max(year),". The overall trend is ", trend, 
#                                           "and the last five years are ", status, "the time series mean of ", mn, ".")))
#   return(plain)
# 
# }
```


## Purpose 

1) To make Will's life easier!! Automate and organize summary output for indicators to make it easier to update the website text. One-stop shop for SMEs?
2) Automate the process for visualization and plain language. Beginnings of PL Automation.
3) Look at indicator status and trends as a whole across regions and indicator group

## Annual Data Sets
- SST
- CHLA
- Beach Closures

### Summary output

```{r, warning=FALSE, message=FALSE, echo=FALSE}

datalist <- c("SST", "Chla", "beach")

summarydf<- data.frame() #blank df to be populated

for (i in datalist) {
  dat<-summarize_me_for_fun(i, time_step = "annual")
  summarydf<-summarydf %>% rbind(dat)
}

DT::datatable(summarydf)
```



```{r, warning=FALSE, message=TRUE, echo=FALSE}

for (i in datalist) {
  plot_me_for_fun(i, time_step = "annual", save = TRUE)
}
```

### Monthly Summary output
## Monthly Data Sets
- AMO
- MEI

**Something wrong in delta calc**

```{r, warning=FALSE, message=FALSE, echo=FALSE}

datalist <- c("amo", "mei")

summarydf<- data.frame() #blank df to be populated

for (i in datalist) {
  dat<-summarize_me_for_fun(i, time_step = "monthly")
  summarydf<-summarydf %>% rbind(dat)
}

DT::datatable(summarydf)
```

```{r, warning=FALSE, message=TRUE, echo=FALSE}

for (i in datalist) {
  plot_me_for_fun(i, time_step = "monthly", save = TRUE)
}
```

## Concerns
Data with gaps in the end of the time series. How to handle 3 or more NAs in the last 5 years??
- Gulf of Alaska Marine Species Distribution
- msddep
- msdlat

```{r, warning=FALSE, message=TRUE, echo=FALSE}
datalist2 <- c("msddep")

for (i in datalist2) {
  plot_me_for_fun(i, save = TRUE)
}
```


## Next Steps
- Link to automated data steps
- Link to Technical Documentation

