# Summary Stats Automation

<!-- # ```{r} -->
<!-- # # Files to use later in rmd -->
<!-- # eiwg_files <- c("SST.csv", "Chla.csv", "beach.csv", "msddep.csv", "msdlat.csv") -->
<!-- #  -->
<!-- # for(i in eiwg_files) { ##### BREAKS on GLPopulation.CSV #################### -->
<!-- #  -->
<!-- #   name <- gsub(".csv","",i) -->
<!-- #    -->
<!-- #   header_read<- read.csv(here::here("data-eiwg",i), nrows = 1) # grabs just the two rows at the top -->
<!-- #   title = colnames(header_read[1]) # title from column names -->
<!-- #   y_lab = as.character(header_read[1,][1]) #ylab from row 1  -->
<!-- #    -->
<!-- #  -->
<!-- #   dat<- assign(name, -->
<!-- #          read.csv(here::here("data-eiwg",i),skip = 2)) -->
<!-- #     colnames(dat) <- tolower(colnames(dat)) -->
<!-- #     dat<- dat %>%  -->
<!-- #     tidyr::pivot_longer(!year, names_to = "region", values_to = "value") %>%  -->
<!-- #     dplyr::mutate(title = c(title), # adds columns with title and y label for figures -->
<!-- #                    y_lab = c(y_lab)) -->
<!-- #    -->
<!-- #   write.csv(dat, file = here::here(paste0("data-raw/", name, ".csv"))) #writes data to data-raw to be pull directly from functions below -->
<!-- # } -->
<!-- #  -->
<!-- #  -->
<!-- # ``` -->


```{r setup, include=FALSE, warning=FALSE, message=TRUE, echo=FALSE}

library(tidyverse)

shade.alpha <- 0.3
shade.fill <- "lightgrey"
lwd <- 1
pcex <- 2
trend.alpha <- 0.5
trend.size <- 2
hline.size <- 1
hline.alpha <- 0.35
hline.lty <- "dashed"
label.size <- 5
hjust.label <- 1.5
letter_size <- 4
ylim_re <- c(5e6, 30e6)
ylim_rd <- c(1.75,2.75)
ylim_ra  <- c(0, 2e6)

#issues: Can't handle 5+ NAs, doesn't do monthly
summarize_me_for_fun<- function(data){
  dat_all<- read.csv(here::here("data-raw", paste0(i, ".csv")))

  dat_all2 <- dat_all %>% 
    dplyr::group_by(region, title, y_lab) %>% 
    dplyr::summarise(mn = mean(value, na.rm = TRUE)) 
  
  sum_dat <- dat_all %>% 
    dplyr::group_by(region) %>% 
    dplyr::filter(year %in% c(max(year):(max(year) - 4))) %>% 
    dplyr::mutate(mn5 = mean(value, na.rm=TRUE)) %>% 
    dplyr::left_join(dat_all2, by = c("region", "title", "y_lab")) %>% 
    dplyr::summarise(status = dplyr::case_when(mn5>mn ~ "greater than", 
                                               mn5<mn ~ "less than"), 
                     mn5 = mn5)
  
  summary_df<- dat_all %>% 
    dplyr::group_by(region) %>% 
    dplyr::left_join(sum_dat, by = c("region")) %>%
    dplyr::mutate(#mn5 = mean(value, na.rm=TRUE),
                  delta = stats::predict(lm(year~value))[length(stats::predict(lm(year~value)))] - stats::predict(lm(year~value))[1], 
                  Z = abs(delta)-(sd(value, na.rm=TRUE)), 
                  trend = ifelse(Z > 0, paste0("significantly"), paste0("")), 
                  gauge = round(stats::ecdf(value)(mn5) *100)) %>% 
    
    dplyr::left_join(dat_all2, by = c("region", "title" ,"y_lab")) %>% 
    dplyr::summarise(indicator = unique(title), 
                     region = unique(region), 
                     #year = unique(year),
                     mn = unique(mn), 
                     mn5 = unique(mn5),
                     status = unique(status), 
                     trend = unique(trend), 
                     delta = unique(delta), 
                     Z = unique(Z),
                     gauge = unique(gauge))  
    
  
  return(summary_df)
}
#issues: Can't handle 5+ NAs, doesn't do monthly

plot_me_for_fun<- function(data, save = TRUE){
    dat <- read.csv(here::here("data-raw", paste0(i, ".csv")))
    
    values<- dat %>% 
      group_by(region) %>% 
      dplyr::mutate(mn  = mean(value), 
                    upper = mn + sd(value), 
                    lower = mn - sd(value))
    pt<-dat %>% 
     dplyr::left_join(values, by = c("region", "value", "title", "y_lab", "year")) %>% 
      ggplot2::ggplot(aes(x = year, y = value, group = region))+
      ggplot2::annotate("rect", fill = "green", alpha = 0.3,
                        xmin = max(dat$year - 5), xmax = max(dat$year),
                        ymin = -Inf, ymax = Inf)+
      theme(panel.background = element_rect(fill = "transparent"), # bg of the panel
            panel.border = element_rect(colour = "black", fill=NA, size=1)) +
      ggplot2::geom_line(color = "red")+
      ggplot2::geom_point(color = "red")+
      ggplot2::geom_hline(aes(yintercept = mn),
                 size = hline.size,
                 alpha = hline.alpha,
                 linetype = hline.lty) +
      #geom_hline(aes(yintercept = sd_TEMP,),
      ggplot2::geom_hline(aes(yintercept = upper),
                 size = hline.size,
                 alpha = hline.alpha,
                 linetype = "dotted") +
      #geom_hline(aes(yintercept = -sd_TEMP,),
      ggplot2::geom_hline(aes(yintercept = lower),
                 size = hline.size,
                 alpha = hline.alpha,
                 linetype = "dotted") +
      ggplot2::ggtitle(paste(values$title))+
      ggplot2::ylab(paste(values$y_lab))+ ## Fix this with Will
      ggplot2::xlab(expression("Year"))+
      ggplot2::facet_wrap(~region, scales = "free")
    print(pt)
  #   if(save == TRUE){
  #     png(here::here(paste0("plots_saved/", something, ".png")))
  # }else{ }
}

# plain_language<- function(data){
#   plain<-  summarydf %>% 
#     group_by(region, title) %>% 
#     dplyr::mutate(PlainLanguage = c(paste("this figure shows", title, " in ", y_lab, "ranging from ", min(value), "to", max(value), 
#                                           "from", min(year), "to ", max(year),". The overall trend is ", trend, 
#                                           "and the last five years are ", status, "the time series mean of ", mn, ".")))
#   return(plain)
# 
# }
```


## Purpose 

1) To make Will's life easier!! Automate and organize summary output for indicators to make it easier to update the website text.
2) Automate the process for the Drupal writers? is this right
3) Potentially look at indictaor status and trends as a whole across regions and indicator group? Language?

## Data Sets
- SST
- CHLA
- Beach Closures

### Summary output
```{r, warning=FALSE, message=FALSE, echo=FALSE}

datalist <- c("SST", "Chla", "beach")

summarydf<- data.frame() #blank df to be populated

for (i in datalist) {
  dat<-summarize_me_for_fun(i)
  summarydf<-summarydf %>% rbind(dat)
}

DT::datatable(summarydf)


```



```{r, warning=FALSE, message=TRUE, echo=FALSE}


for (i in datalist) {
  plot_me_for_fun(i, save = TRUE)
}
```


## Concerns
Data with gaps in the end of the time series. How to handle 3 or more NAs in the last 5 years??
- Gulf of Alaska Marine Species Distribution
- msddep
- msdlat

```{r, warning=FALSE, message=TRUE, echo=FALSE}
datalist2 <- c("msddep")

for (i in datalist2) {
  plot_me_for_fun(i, save = TRUE)
}
```


## Next Steps
- Link to automated data steps
- Link to Technical Documentation

