---
title: "Summary Playground"
author: "Kimberly Bastille, Modified by Will Klajbor"
date: "5/17/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

shade.alpha <- 0.3
shade.fill <- "lightgrey"
lwd <- 1
pcex <- 2
trend.alpha <- 0.5
trend.size <- 2
hline.size <- 1
hline.alpha <- 0.35
hline.lty <- "dashed"
label.size <- 5
hjust.label <- 1.5
letter_size <- 4
ylim_re <- c(5e6, 30e6)
ylim_rd <- c(1.75,2.75)
ylim_ra  <- c(0, 2e6)


```

## Read in Data
Includes all annual NaMES Indicators for now
```{r read-in-data}
library(tidyr)
library(here)
library(dplyr)

#loop in EIWG Data

eiwg_files <- list.files(here::here("data-eiwg"))
for(i in eiwg_files) {
  name <- gsub(".csv","",i)
  assign(name,
  read.csv(here("data-eiwg",i), skip = 2))
}

#list of indicator names

indicator_names <- gsub(".csv","",eiwg_files)
dfs <- Filter(function(x) is(x, "data.frame"), mget(ls()))
res<- lapply(dfs, function(w) {gather(w, region, unit, 2:ncol(w), factor_key=TRUE)})
indicators <- mapply(cbind, res, "title"=indicator_names, SIMPLIFY=F)

#produces list of indicator dfs in same format as dummy data
```

##Old Code
For Dummy Data
```{r data-list}
#datalist<-c("dataA", "dataB", "dataC", "dataD", "dataE")
#datalist<-c(indicators)

# just use "indicators"
```

```{r old functions-to-write-into-package}
# 
# summarize_me_for_fun<- function(data){
#   
#   #dat_all<- read.csv(here::here("data-eiwg", paste0(data, ".csv"))) 
#  dat_all <- i %>% 
#     dplyr::group_by(region) %>% 
#     dplyr::mutate(mn = mean(y, na.rm = TRUE))
#   
#   summary_df<-read.csv(here::here("data-eiwg", paste0(data, ".csv"))) %>% 
#     dplyr::group_by(region) %>% 
#     dplyr::filter(x %in% c(max(x):(max(x) - 4))) %>% 
#     dplyr::mutate(mn5 = mean(y, na.rm=TRUE), 
#                   status = ifelse(mn5>unique(dat_all$mn), paste0("greater than"), paste0("less than")), 
#                   delta = stats::predict(lm(x~y))[length(stats::predict(lm(x~y)))] - stats::predict(lm(x~y))[1], 
#                   Z = abs(delta)-(sd(dat_all$y, na.rm=TRUE)), 
#                   trend = ifelse(Z > 0, paste0("significantly"), paste0("")), 
#                   guage = round(stats::ecdf(dat_all$y)(mn5) *100)) %>% 
#     dplyr::summarise(title = unique(title), 
#                      region = unique(region), 
#                      mn = unique(dat_all$mn), 
#                      mn5 = unique(mn5),
#                      status = unique(status), 
#                      trend = unique(trend), 
#                      delta = unique(delta), 
#                      Z = unique(Z),
#                      guage = unique(guage))
# 
#   return(summary_df)
# }
# 
# 
# 
# plot_me_for_fun<- function(data){
#  # dat<-read.csv(here::here("data-eiwg", paste0(data, ".csv")))
#   dat <- i
#   pt<-dat %>% ggplot2::ggplot(aes(x = x, y = y))+
#     ggplot2::annotate("rect", fill = "green", alpha = 0.3,
#                       xmin = max(dat$x - 10), xmax = max(dat$x),
#                       ymin = -Inf, ymax = Inf)+
#     theme(panel.background = element_rect(fill = "transparent"), # bg of the panel
#           panel.border = element_rect(colour = "black", fill=NA, size=1)) + 
#     ggplot2::geom_line(color = "red")+
#     ggplot2::geom_point(color = "red")+
#     ggplot2::geom_hline(aes(yintercept = mean(y),),
#                size = hline.size,
#                alpha = hline.alpha,
#                linetype = hline.lty) +
#     #geom_hline(aes(yintercept = sd_TEMP,),
#     ggplot2::geom_hline(aes(yintercept = mean(y) + sd(y),),
#                size = hline.size,
#                alpha = hline.alpha,
#                linetype = "dotted") +
#     #geom_hline(aes(yintercept = -sd_TEMP,),
#     ggplot2::geom_hline(aes(yintercept = mean(y)-sd(y),),
#                size = hline.size,
#                alpha = hline.alpha,
#                linetype = "dotted") +
#     ggplot2::ggtitle(paste(dat$title))+
#     ggplot2::ylab(expression("fix this"))+ ## Fix this with Will
#     ggplot2::xlab(expression("Year"))+
#     ggplot2::facet_wrap(~region)
#   print(pt)
# }
```

## New Code
Adjusted for variable names on EIWG Data
```{r new-functions}

#issues: Can't handle 5+ NAs, doesn't do monthly
summarize_me_for_fun<- function(data){
  
  dat_all <- i %>% 
    dplyr::group_by(region) %>% 
    dplyr::mutate(mn = mean(unit, na.rm = TRUE))
  
  summary_df <- i %>% 
    dplyr::group_by(region) %>% 
    dplyr::filter(Year %in% c(max(Year):(max(Year) - 4))) %>% 
    dplyr::mutate(mn5 = mean(unit, na.rm=TRUE), 
                  status = ifelse(mn5>unique(mean(dat_all$mn)), paste0("greater than"), paste0("less than")), 
                  delta = stats::predict(lm(Year~unit))[length(stats::predict(lm(Year~unit)))] - stats::predict(lm(Year~unit))[1], 
                  Z = abs(delta)-(sd(dat_all$unit, na.rm=TRUE)), 
                  trend = ifelse(Z > 0, paste0("significantly"), paste0("")), 
                  gauge = round(stats::ecdf(dat_all$unit)(mn5) *100)) %>% 
    dplyr::summarise(indicator = unique(title), 
                     region = unique(region), 
                     year = unique(Year),
                     mn = unique(mean(dat_all$mn)), 
                     mn5 = unique(mn5),
                     status = unique(status), 
                     trend = unique(trend), 
                     delta = unique(delta), 
                     Z = unique(Z),
                     gauge = unique(gauge))
  
  return(summary_df)
}
#issues: Can't handle 5+ NAs, doesn't do monthly

plot_me_for_fun<- function(data){
    dat <- i
    pt<-dat %>% ggplot2::ggplot(aes(x = Year, y = unit))+
      ggplot2::annotate("rect", fill = "green", alpha = 0.3,
                        xmin = max(dat$Year - 5), xmax = max(dat$Year),
                        ymin = -Inf, ymax = Inf)+
      theme(panel.background = element_rect(fill = "transparent"), # bg of the panel
            panel.border = element_rect(colour = "black", fill=NA, size=1)) +
      ggplot2::geom_line(color = "red")+
      ggplot2::geom_point(color = "red")+
      ggplot2::geom_hline(aes(yintercept = mean(unit),),
                 size = hline.size,
                 alpha = hline.alpha,
                 linetype = hline.lty) +
      #geom_hline(aes(yintercept = sd_TEMP,),
      ggplot2::geom_hline(aes(yintercept = mean(unit) + sd(unit),),
                 size = hline.size,
                 alpha = hline.alpha,
                 linetype = "dotted") +
      #geom_hline(aes(yintercept = -sd_TEMP,),
      ggplot2::geom_hline(aes(yintercept = mean(unit)-sd(unit),),
                 size = hline.size,
                 alpha = hline.alpha,
                 linetype = "dotted") +
      ggplot2::ggtitle(paste(dat$title))+
      ggplot2::ylab(expression("fix this"))+ ## Fix this with Will
      ggplot2::xlab(expression("Year"))+
      ggplot2::facet_wrap(~region)
    print(pt)
  }

```


## Plots

These are all the plots you build - look how pretty. 

```{r plots, warning=FALSE}
for (i in indicators) {
  plot_me_for_fun(i)
}
```

## Summary data

```{r summary, warning=FALSE}
summarydf<- data.frame() #blank df to be populated

for (i in indicators) {
  dat<-summarize_me_for_fun(i)
  summarydf<-summarydf %>% rbind(dat)
}

DT::datatable(summarydf)

write.csv(summarydf, file = here::here("data","summarydf.csv"))
```

### Will updates 5/17/23
- Plots and table work, now need to clean up
- Need to find a way to include data with lots of NAs
-- summary table function stops when 5+ Nas are present at end of time series
- Can we find way to plug in monthly data?

### Will's spitballs 5/11/23
- Need to find way to batch read in and reformat existing EIWG data
```{r}
# library(tidyr)
# library(here)
# library(dplyr)
# 
# #test data
# #load in data, skip 2 rows (y label, title)
# SST <- read_csv("data-eiwg/SST.csv", skip = 2)
# 
# SST_long <- gather(SST, region, degC, 2:10, factor_key=TRUE)
# SST_format <- SST_long %>% mutate(title = "SST")
# colnames(SST_format) <- c('y','region','x','title')
# head(SST_format)
# 
# #test batch
# #take a swing at loop
# 
# eiwg_files <- list.files(here::here("data-eiwg"))
# for(i in eiwg_files) {                        
#   name <- gsub(".csv","",i)
#   assign(name,                                   
#   read.csv(here("data-eiwg",i), skip = 2))
# }
# #this^ works but you lose units
# 
# indicator_names <- gsub(".csv","",eiwg_files)
# dfs <- Filter(function(x) is(x, "data.frame"), mget(ls()))
# res<- lapply(dfs, function(w) {gather(w, region, unit, 2:ncol(w), factor_key=TRUE)})
# indicators <- mapply(cbind, res, "title"=indicator_names, SIMPLIFY=F)
# 
# #produces list of indicator dfs in same format as dummy data
```

### Notes for improvement 3/29/23

- Guage read out - Percentile rank of last 5 compared to mean of whole series. 

```{r}
# mn5 = mean(Y5, na.rm=TRUE)            # mean over the eval period
#   mn = mean(co)                         # mean over entire data series
#  # pTileRankNew <- round(((length(co_all[co_all <= mn5])/(length(co_all)+1))*100)) # older version of percentile
#   cdf <- ecdf(co_all)
#  pTileRankECDF <- round(cdf(mn5) *100)
```

- Trend - Is mean of last 5 years above or below 1 SD?

```{r}
# if (trendAnalysis==T) {
#     par(mar=c(2.5,0,3,0))
#     plot(1, xlim=c(0.94,1.06), ylim=c(0.6, 1.6), col=0, axes=F, xlab="", ylab="")
# 
#     # Mean of eval period outside 10th or 90th percentile?
#     points(1, 1.225, pch=20, cex=5)
#     maxNA <- length(Y5) * 0.4 # scale the maximum allowable NAs in eval period
#     print(paste("Max Na's for eval period= ",maxNA))
#     if (sum(is.na(Y5)) < maxNA)  {
#       if (mn5 >= ptile[3])  { text(1, 1.225, col="white", "+", cex=2.6, font=2) }
#       if (mn5 >= ptile[3])  { text(1, 1.225, col="white", "+", cex=2.6, font=2) }
#       if (mn5 <= ptile[1])  { text(1, 1.225, col="white", "-", cex=2.6, font=2) }
# }}
#       # Change over last 5 yrs > 1 s.d.?
#       m1 = lm(Y5~X5)
#       s1 <- summary(m1)
#       b1 <- s1$coefficients[2,1]
#       pval <- s1$coefficients[2,4]
#       pred = predict(m1)
#       delta = pred[length(pred)] - pred[1]  # gives magnitude and direction of change over eval period
#       Z = abs(delta)-(sd(co_all, na.rm=T))
# Z positive means the total change over eval period exceeds 1 s.d. of entire series
```

Z if pos use significantly
delta says if it above or below mean


LME column 
Indicator title column
Y-label column
